### 2016-11-18 local-capability-machine.tex
lemma

\usepackage{tikz}
\usepackage{multirow}

@@

\section{Macros}
In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
\begin{description}
\item[\texttt{call f(r)}] Moves the argument to register $\texttt{r}$ (if there are more arguments, then they are moved to the appropriate registers). Sets up a return capability in $r_0$ (this includes storing existing capabilities, so they can be read upon return). Erases all registers but the argument register(s), $\pcreg$, and the return register. Jumps to $\texttt{f}$.
\item[\texttt{malloc x}] Calls malloc to allocates a piece of memory (of an appropriate size). In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
\item[\texttt{hstore x n}] Fetches the capability referred to with \texttt{x} and stores \texttt{n} to it.
\item[\texttt{assert x y}] Fetches \texttt{x} and \texttt{y} and compares them. If they are equal, then execution continues. If they are unequal, then the assertion flag is set to 1 and execution halts.
\end{description}

@@

\begin{figure}
  \label{fig:stack-after-call}
  \centering
  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$r_{\var{stk}} \rightarrow$   & local state\\
   & $\vdots$
\end{tabular}
\caption{Stack after everything has been restored}
\end{figure}
\subsection{Labels}
\texttt{l:} defines a meta level label.

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1. The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}). Notice that an assertion will never terminate in a $\failed$ state.
\subsubsection{Simple use of local capabilities}
Old ML-like program:
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv();
          assert (l == 1)
\end{verbatim}
Assembly program not using stack
\begin{verbatim}
f1: malloc l
    hstore l 1
    call adv()
    assert l 1
    halt
1f:
\end{verbatim}

@@

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
g = fun _ => let x = 0 in
               fun f =>
                 x:=0;
                 f();
                 x:=1;
                 f();
                 assert(x==1)
\end{verbatim}
\begin{lemma}\forcenewline
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{g} & \defeq ((\entry,\glob),\start_g,\addrend_g,\start_g)
  \end{align*}
if
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus \hs_{\malloc} \uplus h_f) \step[n] (\halted,h')
  \]
where 
\begin{align*}
  &\dom(hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(hs_g) = [\start_g,\addrend_g] \\
  &\dom(hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\adv} \uplus \hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}, 1 \mapsto\iota^{\var{nwl}}_{\start_\adv,\addrend_\adv}]}
\end{align*}
and $hs_\flag(\flag) = 0$ and $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$ as well as an entry capability for $\malloc$, 
then
\[
  h'(\flag) = 0
\]
\end{lemma}

>>>>

\usepackage{tikz}
\usepackage{multirow,bigdelim}

@@

\section{Macros}
In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
\begin{description}
\item[\texttt{call f(r)}] Moves the argument to a register (if there are more arguments, then they are moved to the appropriate registers). Sets up a return capability in $r_0$ (this includes storing existing capabilities, so they can be read upon return). Erases all registers but the argument register(s), $\pcreg$, the $\malloc$ register (if there is one), and the return register. Jumps to $\texttt{f}$.
\item[\texttt{malloc x}] Calls malloc to allocates a piece of memory (of an appropriate size). In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
\item[\texttt{hstore x n}] Fetches the capability referred to with \texttt{x} and stores \texttt{n} to it.
\item[\texttt{assert x y}] Fetches \texttt{x} and \texttt{y} and compares them. If they are equal, then execution continues. If they are unequal, then the assertion flag is set to 1 and execution halts.
\end{description}

@@

\begin{figure}
  \label{fig:stack-after-call}
  \centering
  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$r_{\var{stk}} \rightarrow$   & local state\\
   & $\vdots$
\end{tabular}
\caption{Stack after everything has been restored}
\end{figure}
\subsection{Labels}
\texttt{l:} is a meta level label that can be used to refer to a specific address.

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1. The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}). Notice that an assertion will never terminate in a $\failed$ state.
\subsubsection{Simple use of local capabilities}
ML-like program we had before:
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv();
          assert (l == 1)
\end{verbatim}
Assembly program not using stack
\begin{verbatim}
f1: malloc l
    hstore l 1
    call adv()
    assert l 1
    halt
1f:
\end{verbatim}
\begin{lemma}[Correctness lemma for \texttt{f1}]
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1},\mathtt{1f},\mathtt{f1}) \\
  \end{align*}
if
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus h_{\malloc} \uplus \hs_f) \step[n] (\halted,h')
  \]
where 
\begin{align*}
  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(\hs_{f1}) = [\mathtt{f1},\mathtt{1f}] \\
  &\dom(\hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]}
\end{align*}
and $hs_\flag(\flag) = 0$. $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
then
\[
  h'(\flag) = 0
\]  
\end{lemma}

Assembly program using the stack:
\begin{verbatim}
f2: push 1
    scall adv()
    pop r1
    assert r1 1
    halt
2f:
\end{verbatim}

@@

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
g = fun _ => let x = 0 in
               fun f =>
                 x:=0;
                 f();
                 x:=1;
                 f();
                 assert(x==1)
\end{verbatim}
\lau{ This example relies on \texttt{adv} actually passing on a proper $\malloc$ - is that a fair assumption? The example can be changed to not require this by assuming x and the function is already in the heap. This change does not remove the difficulties in the original program. }
\begin{lemma}\forcenewline
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{g} & \defeq ((\entry,\glob),\start_g,\addrend_g,\start_g)
  \end{align*}
if
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus \hs_{\malloc} \uplus h_f) \step[n] (\halted,h')
  \]
where 
\begin{align*}
  &\dom(hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(hs_g) = [\start_g,\addrend_g] \\
  &\dom(hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\adv} \uplus \hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}, 1 \mapsto\iota^{\var{nwl}}_{\start_\adv,\addrend_\adv}]}
\end{align*}
and $hs_\flag(\flag) = 0$ and $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$ as well as an entry capability for $\malloc$, 
then
\[
  h'(\flag) = 0
\]
\end{lemma}
